---
title: "OOPA Risk of Material Breach Randomization"
author: "Nathaniel Olin"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(randomizr)

set.seed(100)
```

# Load data

Load full list of BRTs and address data.

```{r}
full <- read_csv(file.path('..', 'data', 'parcels.csv'))
# Convert BRT to integer to drop leading zeroes, for matching
table(nchar(full$brt))
full$brt <- as.integer(full$brt)
table(nchar(full$brt))
```

Load OOPA case-level data

```{r}
oopa <- read_csv(file.path('..', 'data', 'OOPA_report_clean.csv'))
```

## Data quality checks

BRT number should be unique

```{r}
stopifnot(!duplicated(full$brt))
```

Case should be unique on OOPA file

```{r}
stopifnot(!duplicated(oopa$case))
```

## Merge

```{r}
c('BRT rows' = nrow(full),
  'BRT unique' = length(unique(full$brt)),
  'OOPA rows' = nrow(oopa),
  'OOPA unique' = length(unique(oopa$brt)))

dat <- full_join(oopa, full, by = 'brt')
c('Final rows' = nrow(dat),
  'Final unique BRT' = length(unique(dat$brt)))
```

```{r}
c('Final dataset' = nrow(dat),
  'In both datasets' = sum(dat$brt %in% oopa$brt & dat$brt %in% full$brt),
  'In OOPA not BRT (0?)' = sum(dat$brt %in% oopa$brt & ! dat$brt %in% full$brt),
  'In BRT not OOPA (big)' = sum(dat$brt %in% full$brt & ! dat$brt %in% oopa$brt))

stopifnot(full$brt %in% dat$brt,
          oopa$brt %in% dat$brt)
```

It's OK if the "in OOPA but not BRT" cases are mostly complete.

```{r}
table(
  `Status` = dat$status,
  `In OOPA, not BRT` = dat$brt %in% oopa$brt & ! dat$brt %in% full$brt)

dat %>% filter(status == '520 Ongoing' &
                 brt %in% oopa$brt & 
                 ! brt %in% full$brt)
```

```{r}
rm(oopa, full)
```

# Collapse to BRT level

How many BRTs appear more than once?

```{r}
dat %>%
  filter(duplicated(brt)) %>%
  select(brt) %>% unique() %>% nrow()
```

If there are multiple cases with the same BRT:

* If only one is ongoing, take that.
* If none are ongoing, take the most recent one.
* If multiple are ongoing, check the data / report - shouldn't be possible.

```{r}
dat <- dat %>%
  mutate(active = status == '520 Ongoing') %>%
  arrange(brt, desc(active), desc(agree_start)) %>%
  group_by(brt) %>%
  mutate(brt_rank = 1:n()) %>%
  ungroup()

dat %>% 
  filter(brt %in% dat$brt[duplicated(dat$brt)]) %>%
  select(brt, agree_start, brt_rank)
```

```{r}
reject <- subset(dat, !is.na(brt) & brt_rank > 1)

# Did any active OOPAs get kicked out when de-duplicating by BRT?
table(reject$status)
dat[dat$brt %in% reject$brt[reject$status == '520 Ongoing'], ]
# Not enough to matter
```

Deduplicate

```{r}
dat <- subset(dat, is.na(brt) | brt_rank == 1)

stopifnot(!duplicated(dat$brt))
dat$brt_rank <- NULL
dat$active <- NULL
```

# New variables

* Current OOPA status
    * Not in OOPA
    * Currently in OOPA, good standing
    * Currently in OOPA, in material breach
    * Previously in OOPA, completed
    * Previously in OOPA, defaulted
    
```{r}
dat$status[is.na(dat$status)] <- 'Not in OOPA'

# Did people fall behind on payments?
dat$last_date <- as.Date(NA)
dat$last_date[!is.na(dat$default_dt)] <- dat$default_dt[!is.na(dat$default_dt)]
dat$last_date[!is.na(dat$complete_dt)] <- dat$complete_dt[!is.na(dat$complete_dt)]
dat$last_date[is.na(dat$last_date) & !is.na(dat$case)] <- Sys.Date()

dat <- dat %>%
  mutate(months_in_agree = 
           (lubridate::year(last_date) * 12 + lubridate::month(last_date)) -
           (lubridate::year(agree_start) * 12 + lubridate::month(agree_start)),
         num_diff = months_in_agree - num_pmts)

dat$status[dat$status == '520 Ongoing' & 
             dat$num_diff > 1] <- '520 Ongoing, Breached'
```

Add in missing values in stratification variables (or drop when small)

```{r}
dat$tier[is.na(dat$tier)] <- 'Missing'

table(missing_mail = is.na(dat$different_mail), 
      in_oopa = is.na(dat$case))
# Drop these records
dat <- subset(dat, !is.na(different_mail))

table(missing_zip = is.na(dat$property_zip5),
      in_oopa = is.na(dat$case))
# Don't want to drop active OOPA cases so
dat$property_zip5[is.na(dat$property_zip5)] <- 'Missing'
```



# Set up randomization

## Configure parameters

```{r}
# Probability of assignment to treatment
prob <- 0.5
# Condition names
conditions <- c('Control', 'New Letter')
# Blocking variables
block_var <- c('status', 'tier', 'property_zip5', 'different_mail')
```

## Distribution of block variables

```{r}
for(i in seq_along(block_var)){
  p <- dat %>%
    filter(!is.na(case)) %>%
    ggplot(aes_string(block_var[i])) + geom_bar() + coord_flip()
  print(p)
}
```


## Create blocks

```{r}
dat$block <- do.call(paste, c(dat[, block_var], sep = '|'))
```

# Randomize

```{r}
dat$treat <- randomizr::block_ra(blocks = dat$block, 
                                 prob = prob,
                                 conditions = conditions)
```

## Checks

Counts / proportions

```{r}
addmargins(table(dat$block, dat$treat))
prop.table(table(dat$block, dat$treat), margin = 1)
```

Block size

```{r}
dat %>%
  group_by(block) %>%
  summarize(block_size = n()) %>%
  ungroup() %>%
  arrange(block_size) %>%
  mutate(prop = block_size / sum(block_size),
         cumul_prop = cumsum(prop)) %>%
  print()
```

Balance

```{r}
f <- as.formula(sprintf('treat ~ %s', paste(block_var, collapse = ' + ')))
bal_check <- glm(formula = f, data = dat, family = binomial(link = 'logit'))

z <- qnorm(0.975)

broom::tidy(bal_check) %>%
  mutate(ci_low = estimate - z * std.error,
         ci_high = estimate + z * std.error) %>%
  ggplot(aes(term, estimate)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high)) +
  coord_flip()
```

# Export

Full timestamped randomization

```{r}
ts <- format.Date(Sys.time(), format = '%Y-%m-%d_%H-%M-%S')
write_csv(dat, sprintf('full_randomization_%s.csv', ts))
```

Skinny randomization file

```{r}
dat %>%
  select(brt, treat) %>%
  write_csv(file.path('..', 'data', 'rando.csv'))
```

