---
title: "OOPA Risk of Material Breach Randomization"
author: "Nathaniel Olin"
date: "`r date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
---

This script randomly assigns all BRTs in the City of Philadelphia to receive either old letters or new letters (50-50 split), and exports:

* A timestamped file of the randomization for backup purposes,
* A skinny `rando.csv` file with BRT and treatment assignment, for internal use, and
* A timestamped export file for sending to the OIT Revenue office.

```{r knitr-options, include = F}
options(cli.width = 80)
```

```{r}
set.seed(100)
```


```{r libraries}
library(tidyverse)
```

# Load data

```{r read-data}
dat <- read_csv(file.path('..', 'data', 'basefile.csv'))
dat$property_zip5 <- as.character(dat$property_zip5)
```

## Checks

```{r}
# Duplicates
sapply(dat, function(x) sum(duplicated(x)))
stopifnot(!duplicated(dat$brt))

# Missing values
sapply(dat, function(x) sum(is.na(x)))
```

## Cleaning

```{r}
# Bin small zipcodes
dat <- dat %>% 
  group_by(property_zip5) %>%
  mutate(zip_oopa_count = sum(!is.na(case))) %>%
  ungroup() %>%
  mutate(zip_bin = if_else(zip_oopa_count < 25, 'Small Zip', property_zip5)) %>%
  ungroup()

dat %>%
  filter(zip_bin == 'Small Zip') %>%
  count(property_zip5)
```

# Set up randomization

## Configure parameters

```{r rando-parameters}
# Probability of assignment to treatment
prob <- 0.5
# Condition names
conditions <- c('Control', 'New Letter')
# Blocking variables
block_var <- c('status', 'tier', 'different_mail',
               'bin_no_hs', 'bin_college', 'bin_non_english')
```

## Check distribution of block variables {.tabset}

### Currently within OOPA

```{r block-distribution-oopa}
for(i in block_var){
  p <- dat %>%
    filter(!is.na(case)) %>%
    ggplot(aes_string(i)) + 
    geom_bar() +
    labs(title = sprintf('Distribution of %s', i))
  print(p)
}
```

### Full sample

```{r block-distribution-all}
for(i in block_var){
  p <- dat %>%
    ggplot(aes_string(i)) + 
    geom_bar() +
    labs(title = sprintf('Distribution of %s', i))
  print(p)
}
```

## Create blocks

```{r block-create}
dat$block <- do.call(paste, c(dat[, block_var], sep = '|'))
```

# Randomize + quality checks

## Randomization

```{r rando}
dat$treat <- randomizr::block_ra(blocks = dat$block, 
                                 prob = prob,
                                 conditions = conditions)
```

## Treatment assignment

```{r assign-table-all}
tab <- table(dat$status, dat$treat)
tab %>% knitr::kable(format.args = list(big.mark = ','))
tab %>% prop.table(margin = 1) %>% knitr::kable()
```

## Proportion of sample in small blocks {.tabset}

### Currently within OOPA

```{r block-table-oopa}
dat %>%
  filter(!is.na(case)) %>%
  group_by(block) %>%
  summarize(block_size = n()) %>%
  group_by(block_size) %>%
  summarize(
    n_strata = n(),
    n_in_block = sum(block_size)) %>%
  mutate(prop = n_in_block / sum(n_in_block),
         cumul_prop = scales::percent(cumsum(prop))) %>%
  select(cumul_prop, block_size) %>%
  filter(block_size %in% c(1, 2, 3, 5, 10, 20)) %>%
  knitr::kable(col.names = c(
    'Prop. in blocks $\\leq$ block size', 'Block size'),
    format = 'html') %>%
  kableExtra::kable_styling(full_width = F)
```

### Full sample

```{r block-table-all}
dat %>%
  group_by(block) %>%
  summarize(block_size = n()) %>%
  group_by(block_size) %>%
  summarize(
    n_strata = n(),
    n_in_block = sum(block_size)) %>%
  mutate(prop = n_in_block / sum(n_in_block),
         cumul_prop = scales::percent(cumsum(prop))) %>%
  select(cumul_prop, block_size) %>%
  filter(block_size %in% c(1, 2, 3, 5, 10, 20)) %>%
  knitr::kable(col.names = c(
    'Prop. in blocks $\\leq$ block size', 'Block size'),
    format = 'html') %>%
  kableExtra::kable_styling(full_width = F)
```

## Covariate Balance {.tabset}

Set confidence level

```{r ci-level}
conf <- 0.90
z <- qnorm(conf + (1 - conf) / 2)
```

### Currently within OOPA

```{r balance-oopa, fig.width = 5, fig.height = 8}
f <- as.formula(sprintf('treat ~ %s', paste(block_var, collapse = ' + ')))
bal_check <- glm(formula = f, 
                 data = dat[!is.na(dat$case), ], 
                 family = binomial(link = 'logit'))

summary(bal_check)

broom::tidy(bal_check) %>%
  mutate(ci_low = estimate - z * std.error,
         ci_high = estimate + z * std.error) %>%
  ggplot(aes(term, estimate)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high)) +
  geom_hline(yintercept = 0, linetype = 'dashed') +
  coord_flip() +
  theme_bw()
```

### Full sample 

```{r balance-all, fig.width = 5, fig.height = 8}
f <- as.formula(sprintf('treat ~ %s', paste(block_var, collapse = ' + ')))
bal_check <- glm(formula = f, 
                 data = dat, 
                 family = binomial(link = 'logit'))

summary(bal_check)

broom::tidy(bal_check) %>%
  mutate(ci_low = estimate - z * std.error,
         ci_high = estimate + z * std.error) %>%
  ggplot(aes(term, estimate)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high)) +
  geom_hline(yintercept = 0, linetype = 'dashed') +
  coord_flip() +
  theme_bw()
```


# Export

```{r}
ts <- format.Date(Sys.time(), format = '%Y-%m-%d_%H-%M-%S')
```

## Internal

Full timestamped randomization

```{r export-full}
write_csv(dat, file.path('..', 'data', sprintf('full_randomization_%s.csv', ts)))
```

Skinny randomization file

```{r export-skinny}
dat %>% select(brt, treat) %>%
  write_csv(file.path('..', 'data', 'rando.csv'))
```

## External

```{r}
if(! 'export' %in% list.files(file.path('..', 'data')))
  dir.create(file.path('..', 'data', 'export'))
dat %>% select(brt, treat) %>% write_csv(
  file.path('..', 'data', 'export', sprintf('BRT_OOPA-RISK-EXPERIMENT_%s.csv', ts)))
```

