---
title: "OOPA Risk of Material Breach experiment analysis"
author: "Nathaniel Olin"
date: "`r date()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
---

```{r setup, message = F, warning = F}
knitr::opts_chunk$set(fig.width = 5, fig.height = 5)
library(tidyverse)
library(modmarg)
options(width = 80)
```

```{r functions}
treat_table <- function(dat, treatment, group_var){
  treatment <- enquo(treatment)
  group_var <- enquo(group_var)
  dat %>%
    count(!! treatment, !! group_var) %>%
    spread(!! group_var, n, sep = ": ") %>%
    # mutate(Total = reduce(select_if(., is.numeric), `+`)) %>%
    mutate_if(is.numeric, function(x) sprintf(
      "%s (%s)", 
      scales::comma(x), 
      scales::percent(round(x / sum(x), 3))))
}

sub_marg <- function(mod, subgroup, 
                     var_interest = "treat", 
                     cofint = 0.9,
                     ...){
  stopifnot(is.factor(mod$model[[subgroup]]))
  res <- list()
  for(i in levels(mod$model[[subgroup]])) {
    res <- c(
      res, 
      marg(mod, var_interest = var_interest, cofint = cofint,
           data = mod$model[mod$model[[subgroup]] == i, ],
           ...)
    )
  }
  names(res) <- levels(mod$model[[subgroup]])
  res <- bind_rows(res, .id = sprintf("subgroup: %s", subgroup))
  res[[sprintf("subgroup: %s", subgroup)]] <- factor(
    res[[sprintf("subgroup: %s", subgroup)]], 
    levels = levels(mod$model[[subgroup]]))
  
  res
}

mytab <- function(lev, eff){
  lev <- bind_rows(lev)
  eff <- bind_rows(eff)
  
  var_interest <- unique(gsub(" = .*$", "", lev$Label))
  lev$Label <- gsub("^.* = ", "", lev$Label)
  eff$Label <- gsub("^.* = ", "", eff$Label)
  
  res <- lev[, 1:5] %>%
  select(Label, Level = Margin, Standard.Error) %>%
  left_join(eff[, 1:5], by = "Label") %>%
  rename(var_interest = Label,
         Effect = Margin,
         SE.Level = Standard.Error.x,
         SE.Effect = Standard.Error.y)
  
  names(res)[1] <- var_interest
  
  res
}

myplot <- function(res, dat){
  res <- bind_rows(res)
  
  # clean up treatment variable
  var_interest <- unique(gsub(" = .*$", "", res$Label))
  names(res)[names(res) == "Label"] <- var_interest
  res[[var_interest]] <- gsub("^.* = ", "", res[[var_interest]])
  res$Label <- scales::percent(round(res$Margin, 3))
  
  # get n size

  if (any(grepl("^subgroup: ", names(res)))){
    subgroup <- names(res)[grepl("^subgroup: ", names(res))]
    subgroup <- gsub("^subgroup: ", "", subgroup)
    names(res)[grepl("^subgroup: ", names(res))] <- subgroup
    
    res <- dat %>%
      count_(c(var_interest, subgroup)) %>%
      left_join(res, by = c(var_interest, subgroup))
    
    res[[subgroup]] <- sprintf(
      "%s\nn = %s",
      res[[subgroup]],
      scales::comma(res$n))
    
    p <- filter(res, !is.na(P.Value)) %>%
      ggplot(aes_string(
        x = subgroup, y = "Margin",
        fill = var_interest,
        label = "Label",
        ymin = "`Lower CI (90%)`",
        ymax = "`Upper CI (90%)`"))
  } else {
    res <- dat %>%
      count_(var_interest) %>%
      left_join(res, by = var_interest)
    
    res[[var_interest]] <- sprintf(
      "%s\nn = %s",
      res[[var_interest]],
      scales::comma(res$n))
    
    p <- filter(res, !is.na(P.Value)) %>%
      ggplot(aes_string(
        x = var_interest, y = "Margin",
        fill = var_interest,
        label = "Label",
        ymin = "`Lower CI (90%)`",
        ymax = "`Upper CI (90%)`"))
  }
  
  if(any(is.na(res$P.Value))){
    p <- p + 
      scale_y_continuous(labels = scales::percent) +
      geom_hline(yintercept = 0) +
      scale_fill_manual(values = '#f3c613')
  } else {
    p <- p + 
      scale_y_continuous(labels = scales::percent,
                         expand = c(0, 0, 0.05, 0)) +
      scale_fill_manual(values = c('#2176d2', '#f3c613'))
  }
  
  p <- p + 
    geom_col(show.legend = F) +
    geom_errorbar(width = 0.2) +
    geom_label(fill = "white") + #aes(vjust = ifelse(Margin > 0, -1, 1))) +
    theme_bw() +
    theme(text = element_text(size = 16),
          axis.text = element_text(color = "black")) +
    labs(x = "")
  p
}

psr <- function(mod) {
  # Pseudo-R2
  mod_null <- update(mod, ~ 1)

  #get log likelihoods for both models
  ll_full <- as.numeric(logLik(mod))
  ll_null <- as.numeric(logLik(mod_null))

  # get rank for full model to use in adjusted R2
  rank_full <- mod$rank

  # check that rank for null model is 1, otherwise throw an error
  if(mod_null$rank != 1){
    stop("null model created incorrectly")
  }

  # calculate mcfaddens r2 and adjusted r2
  psr <- c(
    "mcfaddens.R2" = 1 - (ll_full / ll_null), 
    "mcfaddens.adj.R2" = 1 - ((ll_full - rank_full) / ll_null)
  ) 
  
  print(psr)

}
```

```{r data, warning = F, message = F}
# Load Data

# Randomization
rando <- read_csv(file.path("..", "data", "rando.csv"))

# Covariates
base <- read_csv(file.path("..", "data", "basefile.csv")) %>%
  mutate(previously_defaulted = factor(ifelse(
    !is.na(default_dt),
    "Defaulted before",
    "Never defaulted"))) %>%
  select(brt, status, agree_start, tier,
         property_zip5, different_mail,
         pct_college, pct_no_hs, pct_non_english,
         bin_no_hs, bin_college, bin_non_english,
         previously_defaulted)

# Unierse data
letters <- file.path("..", "data", "mailings") %>%
  list.files(full.names = T) %>%
  lapply(read_csv) %>%
  bind_rows()
names(letters) <- tolower(names(letters))
letters <- letters %>%
  mutate(brt = as.integer(opa_number),
         case = as.integer(case_number),
         risk_date = lubridate::mdy(due_date)) %>%
  select(case, brt, risk_date) %>%
  filter(!duplicated(.))

# Outcome data
def <- read_csv(file.path("~", "projects", "oopa-monthly-reports", "data", "OOPA_default.csv"),
                col_types = list(opa_number = "i")) %>%
  rename(brt = opa_number, default_date = date) %>%
  select(brt, default_date) %>%
  filter(default_date > "2018-05-01")

breach <- file.path("~", "projects", "oopa-monthly-reports", "data", "OOPA_material_breach.csv") %>%
  read_csv() %>%
  filter(!is.na(notice)) %>%
  rename(breach_date = report_date) %>%
  select(case, breach_date) %>%
  filter(breach_date > "2018-05-01") %>%
  arrange(breach_date) %>%
  filter(!duplicated(case))
```

```{r}
dat <- left_join(letters, rando, by = "brt") %>%
  left_join(base, by = "brt") %>%
  left_join(def, by = "brt") %>%
  left_join(breach, by = "case") %>%
  mutate(defaulted = !is.na(default_date),
         breached = !is.na(breach_date),
         tier = factor(tier, levels = c("4", "3", "2", "1", "Missing")))

# 3 BRTs have multiple case numbers. 
# If any case breached, consider BRT breached
dat <- dat %>% 
  arrange(breach_date) %>% 
  filter(!duplicated(brt)) 

# Check for duplicates
stopifnot(!any(duplicated(dat$brt)))
stopifnot(!any(duplicated(dat$case)))
```

```{r}
rm(rando, letters, base, def, breach)
```

# Descriptive results

## Outcomes

```{r}
treat_table(dat, treat, breached) %>%
  knitr::kable(align = "r")

treat_table(dat, treat, defaulted) %>%
  knitr::kable(align = "r")
```

# Balance checks

```{r}
covar <- c("tier", "bin_college", "bin_no_hs", "bin_non_english", "different_mail")
```

# Main Analysis

## Material Breach Rates {.tabset}

```{r}
f <- as.formula(sprintf("breached ~ treat + %s", paste(covar, collapse = " + ")))
mod <- glm(f, family = binomial(link = "logit"), data = dat)
lev <- marg(mod, var_interest = "treat", cofint = 0.9)
eff <- marg(mod, var_interest = "treat", type = "effects", cofint = 0.9)
```

### Level plot

```{r}
myplot(lev, mod$model) +
  labs(title = "Material Breach",
       y = "Material Breach Rate")
```

### Effect plot

```{r}
myplot(eff, mod$model) +
  labs(title = "Material Breach",
       y = "Effect of Letter on Material Breach Rate")
```

### Table

```{r}
mytab(lev, eff) %>% knitr::kable()
```

### Regression

```{r}
f
broom::tidy(mod) %>%
  as_data_frame() %>%
  mutate(p.value = round(p.value, 3))
broom::glance(mod)
psr(mod)
```

## Default Rates {.tabset}

```{r}
f <- as.formula(sprintf("defaulted ~ treat + %s", paste(covar, collapse = " + ")))
mod <- glm(f, family = binomial(link = "logit"), data = dat)
lev <- marg(mod, var_interest = "treat", cofint = 0.9)
eff <- marg(mod, var_interest = "treat", type = "effects", cofint = 0.9)
```

### Level plot

```{r}
myplot(lev, mod$model) +
  labs(title = "Defaults",
       y = "Default Rate")
```

### Effect plot

```{r}
myplot(eff, mod$model) +
  labs(title = "Defaults",
       y = "Effect of Letter on Default Rate")
```

### Table

```{r}
mytab(lev, eff) %>% knitr::kable()
```

### Regression

```{r}
f
broom::tidy(mod) %>%
  as_data_frame() %>%
  mutate(p.value = round(p.value, 3))
broom::glance(mod)
psr(mod)
```

# Subgroup Analysis

## Tier

```{r}
subgroup <- "tier"
```

### Material Breach rates {.tabset}

```{r}
f <- as.formula(sprintf(
  "breached ~ treat * %s + %s",
  subgroup, 
  paste(covar[covar != subgroup], collapse = " + ")))

mod <- glm(f, family = binomial(link = "logit"), data = dat)
lev <- sub_marg(mod, subgroup, type = "levels")
eff <- sub_marg(mod, subgroup, type = "effects")
```

#### Plot

```{r}
myplot(eff, mod$model) + labs(
  title = "Change in Material Breach rate\nby OOPA tier", 
  x = "OOPA Tier",
  y = "Effect of New Letter on Material Breach Rate")
```

#### Table

```{r}
mytab(lev, eff) %>% knitr::kable()
```

#### Regression

```{r}
f
broom::tidy(mod) %>%
  as_data_frame() %>%
  mutate(p.value = round(p.value, 3))
broom::glance(mod)
psr(mod)
```


### Default rates {.tabset}

```{r}
f <- as.formula(sprintf(
  "defaulted ~ treat * %s + %s",
  subgroup, 
  paste(covar[covar != subgroup], collapse = " + ")))

mod <- glm(f, family = binomial(link = "logit"), data = dat)
lev <- sub_marg(mod, subgroup, type = "levels")
eff <- sub_marg(mod, subgroup, type = "effects")
```

#### Plot

```{r}
myplot(eff, mod$model) + labs(
  title = "Change in default rate\nby OOPA tier", 
  x = "OOPA Tier",
  y = "Effect of New Letter on Default Rate")
```

#### Table

```{r}
mytab(lev, eff) %>% knitr::kable()
```

#### Regression

```{r}
f
broom::tidy(mod) %>%
  as_data_frame() %>%
  mutate(p.value = round(p.value, 3))
broom::glance(mod)
psr(mod)
```

## Previously Defaulted

```{r}
subgroup <- "previously_defaulted"
```

### Material Breach rates {.tabset}

```{r}
f <- as.formula(sprintf(
  "breached ~ treat * %s + %s",
  subgroup, 
  paste(covar[covar != subgroup], collapse = " + ")))

mod <- glm(f, family = binomial(link = "logit"), data = dat)
lev <- sub_marg(mod, subgroup, type = "levels")
eff <- sub_marg(mod, subgroup, type = "effects")
```

#### Plot

```{r}
myplot(eff, mod$model) + labs(
  title = "Change in Material Breach rate\nby OOPA tier", 
  x = "OOPA Tier",
  y = "Effect of New Letter on Material Breach Rate")
```

#### Table

```{r}
mytab(lev, eff) %>% knitr::kable()
```

#### Regression

```{r}
f
broom::tidy(mod) %>%
  as_data_frame() %>%
  mutate(p.value = round(p.value, 3))
broom::glance(mod)
psr(mod)
```

### Default rates {.tabset}

```{r}
f <- as.formula(sprintf(
  "defaulted ~ treat * %s + %s",
  subgroup, 
  paste(covar[covar != subgroup], collapse = " + ")))

mod <- glm(f, family = binomial(link = "logit"), data = dat)
lev <- sub_marg(mod, subgroup, type = "levels")
eff <- sub_marg(mod, subgroup, type = "effects")
```

#### Plot

```{r}
myplot(eff, mod$model) + labs(
  title = "Change in default rate\nby OOPA tier", 
  x = "OOPA Tier",
  y = "Effect of New Letter on Default Rate")
```

#### Table

```{r}
mytab(lev, eff) %>% knitr::kable()
```

#### Regression

```{r}
f
broom::tidy(mod) %>%
  as_data_frame() %>%
  mutate(p.value = round(p.value, 3))
broom::glance(mod)
psr(mod)
```


